---
title: "Phenotype QC"
author: "Belinda Cornes"
date: "2020-03-11"
output: html_document
---

This script is to check the distribution of the data and to remove any outliers. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)

#filepathp = '/Users/corneb/Documents/MyJax/CS/Projects/qtl2/Univ_of_Penn_Thaiss/Thaiss/three.batches/Thaiss_workflowr_old/data/heh/'
#pheno_file <- dir(path = filepathp, pattern = "Univ_of_Penn_Thaiss_pheno_all.csv", full.names = TRUE)
#phenos <- read.csv(pheno_file)

#phenos <- subset(phenos, !is.na(Time_min))

library(Hmisc)
library(pastecs)
library(psych)

setwd("/Users/corneb/Documents/MyJax/CS/Projects/qtl2/Univ_of_Penn_Thaiss/Thaiss/three.batches/Thaiss_workflowr_old/data/heh/")

gm_DO112_ssqc <- get(load("/Users/corneb/Documents/MyJax/CS/Projects/qtl2/Univ_of_Penn_Thaiss/Thaiss/three.batches/Thaiss_workflowr_old/data/gm_DO112_ssqc.RData"))

library(qtl2)

#phenos[111,1] 

#rownames(phenos) <- phenos$Mouse.ID



#phenosn <- phenos[ind_ids(gm_DO112_ssqc),]

gm_DO112_ssqc$pheno

phenos <- as.data.frame(gm_DO112_ssqc$pheno)

```

### Phenotype : Time
```{r plot sample distribution for Time}

time <- ggplot(data= phenos, aes(x=phenos$Time_min)) + 
geom_histogram(aes(y=..density..), fill="grey", color="white", position="identity") + 
geom_vline(xintercept=mean(phenos$Time_min),size = 1, col="red") + 
labs(x="Time [mins]") + 
stat_function(fun = dnorm, color="blue", size=1, args=list(mean=mean(phenos$Time_min), sd=sd(phenos$Time_min))) +
ggtitle("Distribution of Time [min]") +
theme(plot.title = element_text(hjust = 0.5))

time

Hmisc::describe(phenos$Time_min) 
pastecs::stat.desc(phenos$Time_min) 
psych::describe(phenos$Time_min) 

```
Given the negative skewness of the data, the intent is to transform by square rooting the reflected data, then check the skewness

```{r transforming Time}

rz.transform <- function(y) {
  rankY=rank(y, ties.method="average", na.last="keep")
  rzT=qnorm(rankY/(length(na.exclude(rankY))+1))
  return(rzT)
}

phenos$Time_min_rnorm <- rz.transform(phenos$Time_min)

Time_min_rnorm <- ggplot(data= phenos, aes(x=phenos$Time_min_rnorm)) + 
geom_histogram(aes(y=..density..), fill="grey", color="white", position="identity") + 
geom_vline(xintercept=mean(phenos$Time_min_rnorm),size = 1, col="red") + 
labs(x="Time [mins]") + 
stat_function(fun = dnorm, color="blue", size=1, args=list(mean=mean(phenos$Time_min_rnorm), sd=sd(phenos$Time_min_rnorm))) +
ggtitle("Distribution of (transformed)Time [min] (rnorm)") +
theme(plot.title = element_text(hjust = 0.5))

Time_min_rnorm

Hmisc::describe(phenos$Time_min_rnorm)  
pastecs::stat.desc(phenos$Time_min_rnorm) 
psych::describe(phenos$Time_min_rnorm) 


phenos$trsfTime <- sqrt(max(phenos$Time_min+1) - phenos$Time_min)

trsfTime <- ggplot(data= phenos, aes(x=phenos$trsfTime)) + 
geom_histogram(aes(y=..density..), fill="grey", color="white", position="identity") + 
geom_vline(xintercept=mean(phenos$trsfTime),size = 1, col="red") + 
labs(x="Time [mins]") + 
stat_function(fun = dnorm, color="blue", size=1, args=list(mean=mean(phenos$trsfTime), sd=sd(phenos$trsfTime))) +
ggtitle("Distribution of (transformed)Time [min]") +
theme(plot.title = element_text(hjust = 0.5))

trsfTime

Hmisc::describe(phenos$trsfTime)  
pastecs::stat.desc(phenos$trsfTime) 
psych::describe(phenos$trsfTime) 

````
Now that the transformation improved the fit of the data, we will check for any outliers

```{r checking outliers Time}
#Working out 3SD from mean

phenos$stdTime_min_rnorm <- scale(phenos$Time_min_rnorm)

phenos$stdTime_min_rnormo <- ifelse(phenos$stdTime_min_rnorm >= 3 | phenos$stdTime_min_rnorm <= -3, 'Outlier','Keep')

table(phenos$stdTime_min_rnormo)


phenos$stdtrsfTime <- scale(phenos$trsfTime)

phenos$stdtrsfTimeo <- ifelse(phenos$stdtrsfTime >= 3 | phenos$stdtrsfTime <= -3, 'Outlier','Keep')

table(phenos$stdtrsfTimeo)

```


No outliers were detected. 


### Phenotype : Distance
```{r plot sample distribution for Distance}

distance <- ggplot(data= phenos, aes(x=phenos$Distance_m)) + 
geom_histogram(aes(y=..density..), fill="grey", color="white", position="identity") + 
geom_vline(xintercept=mean(phenos$Distance_m),size = 1, col="red") + 
labs(x="Distance [m]") + 
stat_function(fun = dnorm, color="blue", size=1, args=list(mean=mean(phenos$Distance_m), sd=sd(phenos$Distance_m))) +
ggtitle("Distribution of Distance [m]") +
theme(plot.title = element_text(hjust = 0.5))

distance

Hmisc::describe(phenos$Distance_m) 
pastecs::stat.desc(phenos$Distance_m) 
psych::describe(phenos$Distance_m)


phenos$Distance_m_rnorm <- rz.transform(phenos$Distance_m)

Distance_m_rnorm <- ggplot(data= phenos, aes(x=phenos$Distance_m_rnorm)) + 
geom_histogram(aes(y=..density..), fill="grey", color="white", position="identity") + 
geom_vline(xintercept=mean(phenos$Distance_m_rnorm),size = 1, col="red") + 
labs(x="Time [mins]") + 
stat_function(fun = dnorm, color="blue", size=1, args=list(mean=mean(phenos$Distance_m_rnorm), sd=sd(phenos$Distance_m_rnorm))) +
ggtitle("Distribution of (transformed) Distance [m] (rnorm)") +
theme(plot.title = element_text(hjust = 0.5))

Distance_m_rnorm

Hmisc::describe(phenos$Distance_m_rnorm)  
pastecs::stat.desc(phenos$Distance_m_rnorm) 
psych::describe(phenos$Distance_m_rnorm) 

```

Transforming the data using the same method as `Time` worsened the negative skew, so we chose to keep the meaure untransformed. 

```{r checking outliers Distance}
#Working out 3SD from mean
phenos$stddist <- scale(phenos$Distance_m)

phenos$stddisto <- ifelse(phenos$stddist  >= 3 | phenos$stddist  <= -3, 'Outlier','Keep')

table(phenos$stddisto)


phenos$stddist.rnorm <- scale(phenos$Distance_m_rnorm)

phenos$stddist.rnormo <- ifelse(phenos$stddist.rnorm  >= 3 | phenos$stddist.rnorm  <= -3, 'Outlier','Keep')

table(phenos$stddist.rnormo)


```

No outliers were detected. 


### Phenotype : Energy
```{r plot sample distribution for Energy}

energy <- ggplot(data= phenos, aes(x=phenos$Energy_J)) + 
geom_histogram(aes(y=..density..), fill="grey", color="white", position="identity") + 
geom_vline(xintercept=mean(phenos$Energy_J),size = 1, col="red") + 
labs(x="Energy [J]") + 
stat_function(fun = dnorm, color="blue", size=1, args=list(mean=mean(phenos$Energy_J), sd=sd(phenos$Energy_J))) +
ggtitle("Distribution of Engery [J]") +
theme(plot.title = element_text(hjust = 0.5))

energy

Hmisc::describe(phenos$Energy_J) 
pastecs::stat.desc(phenos$Energy_J) 
psych::describe(phenos$Energy_J) 


phenos$Energy_J_rnorm <- rz.transform(phenos$Energy_J)

Energy_J_rnorm <- ggplot(data= phenos, aes(x=phenos$Energy_J_rnorm)) + 
geom_histogram(aes(y=..density..), fill="grey", color="white", position="identity") + 
geom_vline(xintercept=mean(phenos$Energy_J_rnorm),size = 1, col="red") + 
labs(x="Time [mins]") + 
stat_function(fun = dnorm, color="blue", size=1, args=list(mean=mean(phenos$Energy_J_rnorm), sd=sd(phenos$Energy_J_rnorm))) +
ggtitle("Distribution of (transformed) Energy J (rnorm") +
theme(plot.title = element_text(hjust = 0.5))

Energy_J_rnorm

Hmisc::describe(phenos$Energy_J_rnorm)  
pastecs::stat.desc(phenos$Energy_J_rnorm) 
psych::describe(phenos$Energy_J_rnorm) 

```

Trait did not need transforming.  Checking for any outliers:

```{r checking outliers Energy}
#Working out 3SD from mean
phenos$stden <- scale(phenos$Energy_J)
phenos$stdeno <- ifelse(phenos$stden  >= 3 | phenos$stden  <= -3, 'Outlier','Keep')

table(phenos$stdeno)

phenos$stden.rnorm <- scale(phenos$Energy_J_rnorm)
phenos$stden.rnormo <- ifelse(phenos$stden.rnorm  >= 3 | phenos$stden.rnorm  <= -3, 'Outlier','Keep')

table(phenos$stden.rnormo)

```

No outliers were detected. 


## Conclusion

+ rnorm all phenotypes


## All Problematic Samples

No outliers were detected. 

```{r saving new pheno file}

phenos$Mouse.ID <- rownames(phenos)

phenosn <- phenos[c("Mouse.ID","Time_min","Distance_m","Energy_J","rsqrt_Time_min","Time_min_rnorm","Distance_m_rnorm","Energy_J_rnorm")]
#phenos$index <- 1:nrow(phenos)

#phenosn <- merge(phenosn_cc, phenos[c("Mouse.ID","index")], by=c("Mouse.ID"), all=T)

#names(phenosn_cc)[2:5] <- c("Time_min","Distance_m","Energy_J","rsqrt_Time_min")
phenosn[is.na(phenosn)] <- ''


#phenosno <- phenosn[order(phenosn$index),]

write.csv(phenosn,"Univ_of_Penn_Thaiss_pheno_all_ssqc_rnorm.csv", row.names=F, quote=F)
```
