---
title: "Phenotype QC"
author: "Belinda Cornes"
date: "2020-03-11"
output: html_document
---

This script is to check the distribution of the data and to remove any outliers. Types of distribution:

+ normal [no correction]
+ positive skewness [possible correction: square root/logarithm]
+ negative skewness [possible correction: reflect & square root/reflect & logarithm]
+ positive kurtosis
+ negative kurtosis 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)

filepathp = '/Users/corneb/Documents/MyJax/CS/Projects/qtl2/Univ_of_Penn_Thaiss/Thaiss/two.batches/data/'
pheno_file <- dir(path = filepathp, pattern = "Uni_of_Penn_Thaiss_pheno.3.csv", full.names = TRUE)
phenos <- read.csv(pheno_file)

phenos_cc <- subset(phenos, !is.na(Time_.min.))

library(Hmisc)
library(pastecs)
library(psych)

setwd("/Users/corneb/Documents/MyJax/CS/Projects/qtl2/Univ_of_Penn_Thaiss/Thaiss/three.batches/Thaiss_workflowr_old/data/heh/")

```

### Phenotype : Time
```{r plot sample distribution for Time}

time <- ggplot(data= phenos_cc, aes(x=phenos_cc$Time_.min.)) + 
geom_histogram(aes(y=..density..), fill="grey", color="white", position="identity") + 
geom_vline(xintercept=mean(phenos_cc$Time_.min.),size = 1, col="red") + 
labs(x="Time [mins]") + 
stat_function(fun = dnorm, color="blue", size=1, args=list(mean=mean(phenos_cc$Time_.min.), sd=sd(phenos_cc$Time_.min.))) +
ggtitle("Distribution of Time [min]") +
theme(plot.title = element_text(hjust = 0.5))

time

Hmisc::describe(phenos$Time_.min.) 
pastecs::stat.desc(phenos$Time_.min.) 
psych::describe(phenos$Time_.min.) 

```
Given the negative skewness of the data, the intent is to transform by square rooting the reflected data, then check the skewness

```{r transforming Time}

phenos_cc$trsfTime <- sqrt(max(phenos_cc$Time_.min.+1) - phenos_cc$Time_.min.)

trsfTime <- ggplot(data= phenos_cc, aes(x=phenos_cc$trsfTime)) + 
geom_histogram(aes(y=..density..), fill="grey", color="white", position="identity") + 
geom_vline(xintercept=mean(phenos_cc$trsfTime),size = 1, col="red") + 
labs(x="Time [mins]") + 
stat_function(fun = dnorm, color="blue", size=1, args=list(mean=mean(phenos_cc$trsfTime), sd=sd(phenos_cc$trsfTime))) +
ggtitle("Distribution of (transformed)Time [min]") +
theme(plot.title = element_text(hjust = 0.5))

trsfTime

Hmisc::describe(phenos_cc$trsfTime)  
pastecs::stat.desc(phenos_cc$trsfTime) 
psych::describe(phenos_cc$trsfTime) 

````
Now that the transformation improved the fit of the data, we will check for any outliers

```{r checking outliers Time}
#Working out 3SD from mean
phenos_cc$stdtrsfTime <- scale(phenos_cc$trsfTime)

phenos_cc$stdtrsfTimeo <- ifelse(phenos_cc$stdtrsfTime >= 3 | phenos_cc$stdtrsfTime <= -3, 'Outlier','Keep')

table(phenos_cc$stdtrsfTimeo)

```


No outliers were detected. 


### Phenotype : Distance
```{r plot sample distribution for Distance}

distance <- ggplot(data= phenos_cc, aes(x=phenos_cc$Distance_.m.)) + 
geom_histogram(aes(y=..density..), fill="grey", color="white", position="identity") + 
geom_vline(xintercept=mean(phenos_cc$Distance_.m.),size = 1, col="red") + 
labs(x="Distance [m]") + 
stat_function(fun = dnorm, color="blue", size=1, args=list(mean=mean(phenos_cc$Distance_.m.), sd=sd(phenos_cc$Distance_.m.))) +
ggtitle("Distribution of Distance [m]") +
theme(plot.title = element_text(hjust = 0.5))

distance

Hmisc::describe(phenos$Distance_.m.) 
pastecs::stat.desc(phenos$Distance_.m.) 
psych::describe(phenos$Distance_.m.) 

```

Transforming the data using the same method as `Time` worsened the negative skew, so we chose to keep the meaure untransformed. 

```{r checking outliers Distance}
#Working out 3SD from mean
phenos_cc$stddist <- scale(phenos_cc$Distance_.m.)

phenos_cc$stddisto <- ifelse(phenos_cc$stddist  >= 3 | phenos_cc$stddist  <= -3, 'Outlier','Keep')

table(phenos_cc$stddisto)

```

No outliers were detected. 


### Phenotype : Energy
```{r plot sample distribution for Energy}

energy <- ggplot(data= phenos_cc, aes(x=phenos_cc$Energy_.J.)) + 
geom_histogram(aes(y=..density..), fill="grey", color="white", position="identity") + 
geom_vline(xintercept=mean(phenos_cc$Energy_.J.),size = 1, col="red") + 
labs(x="Energy [J]") + 
stat_function(fun = dnorm, color="blue", size=1, args=list(mean=mean(phenos_cc$Energy_.J.), sd=sd(phenos_cc$Energy_.J.))) +
ggtitle("Distribution of Engery [J]") +
theme(plot.title = element_text(hjust = 0.5))

energy

Hmisc::describe(phenos$Energy_.J.) 
pastecs::stat.desc(phenos$Energy_.J.) 
psych::describe(phenos$Energy_.J.) 
```

Trait did not need transforming.  Checking for any outliers:

```{r checking outliers Energy}
#Working out 3SD from mean
phenos_cc$stden <- scale(phenos_cc$Energy_.J.)
phenos_cc$stdeno <- ifelse(phenos_cc$stden  >= 3 | phenos_cc$stden  <= -3, 'Outlier','Keep')

table(phenos_cc$stdeno)

```

No outliers were detected. 


## Conclusion

+ `Time_[min]` was transformed using reflect & square root and will be used in all further analysis
+ `Distance_[m]` was *not* transformed
+ `Energy_[J]` was *not* transformed


## All Problematic Samples

No outliers were detected. 

```{r saving new pheno file}

phenosn_cc <- phenos_cc[c("Mouse.ID","Time_.min.","Distance_.m.","Energy_.J.","trsfTime")]
phenos$index <- 1:nrow(phenos)

phenosn <- merge(phenosn_cc, phenos[c("Mouse.ID","index")], by=c("Mouse.ID"), all=T)

names(phenosn_cc)[2:5] <- c("Time_min","Distance_m","Energy_J","rsqrt_Time_min")
phenosn[is.na(phenosn)] <- ''

phenosno <- phenosn[order(phenosn$index),]

write.csv(phenosno[1:5],"Univ_of_Penn_Thaiss_pheno_all.csv", row.names=F, quote=F)
```
